Generalized Tractability

### 2023-06-04 GeneralizedTractability_230529.tex
Updates from Overleaf

\newcommand{\tpmod}[1]{{\;(\operatorname{mod}\;#1)}}

@@

The question of tractability is one of determining how fast the complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  We want to bound the complexity in terms of a simple function of $\varepsilon^{-1}$ and $d$ and identify simple conditions on the singular values that are equivalent to this bound holding. \kachi{In this paper, we investigate such a function that we refer to as the generalized tractability function, denoted by $T$. A proper definition is given in Section \ref{sec:spt}. However, we define a tractability function
\[
 T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\]
which depends on $\varepsilon^{-1}$, $d$ and the parameter $\vp$ respectively.
}

@@

\fred{Point out that $T$ could increase quite rapidly, but this does not affect our theorems.}

\kachi{In Section \ref{sec:spt}, we give the definitions of the considered tractability notions, in particular (strong) tractability. We consider the strong tractability when the information complexity is independent of the dimension, $d$. We then turn to tractability  when the information complexity is polynomial in $d$ and $\varepsilon^{-1}$. In Section \ref{sec:examples} we provide examples of strong(tractability) functions for both the algebraic and exponential cases. We also consider other notions of tractability such as quasi-polynomial tractability, and a generalized notion of weak tractability. In the case of the generalized weak tractability, we give the necessary and sufficient conditions on the notion of exponential tractability in Section \ref{sec:subh}. }

@@

\subsection{Notation and fundamental definitions} \label{sec:not}
Let $T$ be a tractability function defined as follows,
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty),
\end{equation}
A tractability function is a function of simple form that provides an upper bound on the computational complexity of a problem.  Some examples of tractability functions commonly considered are provided in Section \ref{sec:examples}.

@@

For this definition of tractability to make sense we assume that
\begin{subequations} \label{eq:Tconditions}
\begin{equation} \label{eq:Tnondecreasing}
	T \text{ is non-decreasing in all variables,}
\end{equation}
which implies that the problem does not become easier by decreasing the tolerance, $\varepsilon$, or increasing the dimension, $d$. Furthermore, increasing $\vp$ does not decrease $T(\cdot, \cdot, \vp)$.  Since there are an infinite number of positive eigenvalues, it also makes sense to assume that
\begin{equation}
	\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d \in \naturals, \ \vp \in [0,\infty)^s.
\end{equation}
Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit, which we assume will always be positive:
\begin{equation}
	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp) \ge T(0,1,\vzero) > 0.
\end{equation}
\end{subequations}
Finally, there is a technical assumption required for our analysis.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp),   \quad \forall \varepsilon \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in [1,\infty).
\end{equation}

Certain common notions of tractability are covered by the general situation described above.
{\small
\begin{equation*}
	\begin{array}{r@{\quad}c@{\quad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		\text{strong polynomial algebraic} & \max(1,\varepsilon^{-p}) & 1\\
		\text{polynomial algebraic} & \max(1,\varepsilon^{-p})d^{q} & d^{q}\\
		\text{strong polynomial exponential} &  [\max(1,\log(1 + \varepsilon^{-1})]^p & 1\\
		\text{polynomial exponential } &
		[\max(1,\log(1 + \varepsilon^{-1}))]^p  d^{q} & d^{q} \\
        \text{quasi-polynomial algebraic} & 
        \exp\{p(1+\log(\max(1,\varepsilon^{-1}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}\\
        \text{quasi-polynomial exponential} & 
        \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}

Here above, $\vp = (p,q)$ in some cases and $\vp$ is a scalar in other cases. A general reference
for results on algebraic tractability is \cite{NW08}, for exponential tractability we refer to \cite{HKW19} and \cite{KW19}.

@@

For $\varepsilon$ taking on all positive values, $i(\varepsilon,d,\vp)$ takes on all values greater than or equal to $i(\infty,d,\vp)$, so
\[
i\le C_{\vp}\, T(\varepsilon^{-1},d,\vp)+1 \le C_{\vp} T(\lambda_{i,d}^{-1},d,\vp)+1, \qquad \forall i \ge i(\infty,d,\vp),  \ d \in \naturals.
\]
This implies via our technical assumption \eqref{eq:ptauassume} that for any $\tau > 1$,
\begin{gather*}
	K_{\vp,\tau}\,T (\lambda_{i,d}^{-1},d,\tau \vp) \ge
	[T(\lambda_{i,d}^{-1},d, \vp)]^\tau
	\ge
	\left[\frac{(i-1)}{C_{\vp}}\right]^\tau \qquad \forall i \ge i(\infty,d,\vp),  \ d \in \naturals, \\
	 \frac{1}{T (\lambda_{i,d}^{-1},d,\tau \vp)} \le
	\frac{K_{\vp,\tau}\, C_{\vp}^\tau}{(i-1)^\tau}\qquad \forall i \ge \max\{i(\infty,\vp),2\},  \ d \in \naturals.
\end{gather*}

@@

\begin{example}[Algebraic polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-p}\}\, d^q
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where in this case $\vp = (p,q)$.
Then Theorem \ref{thm_main_tract2} yields the  equivalent condition for algebraic polynomial tractability:
\[
 \sup_{d\in\naturals} \sum_{i=L_{(p,q)} d^{q}}^\infty (\min\{1,\lambda_{i,d}^p\}) d^{-q} < \infty \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\]
Furthermore, the set of optimal $(p,q)$ is a subset of
\begin{multline*}
\{(p^*,q^*) : \text{the above condition holds for all }(p,q) \in (p^*,\infty) \times (q^*,\infty) \\
\text{but not for any }(p,q) \in [0,p^*/2) \times [0,q^*/2)\}.
\end{multline*}
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NW08} \fred{same theorem as above??, what about optimality in \cite[Theorem 5.1]{NW08}}.
\end{example}

\begin{example}[Strong exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \left[\max\{1,\log(1+\varepsilon^{-1})\}\right]^p, \quad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]
where $\vp$ is a scalar. Then Theorem \ref{thm_main_strong_tract2} yields the equivalent condition for strong exponential polynomial tractability \[
 \sup_{d\in\naturals} \sum_{i=L_{p} }^\infty (\min\{1,\log(1+\lambda_{i,d}\})^{p}  < \infty,\quad  \text{ for some } p \ge 0, L_{p} \in \naturals
\] In this case the optimal $p$ is the infimum of all $p$ for which the above condition holds.
\end{example}
\begin{example}[Exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)=\left[\max\{1,\log(1+\varepsilon^{-1}\})\right]^p\, d^q, \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where $\vp = (p,q)$. Then Theorem \ref{thm_main_tract2} yields  the equivalent condition for exponential polynomial tractability
\[
 \sup_{d\in\naturals} \sum_{i=L_{(p,q)} d^{q}}^\infty [\min\{1,\log(1+\lambda_{i,d}\})]^p d^{-q} < \infty  \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\]
\end{example}

\begin{example}[Algebraic quasi-polynomial tractability] Here we consider another common notions of tractability.

@@

\noindent Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(1+\log(\max\{1,\varepsilon^{-1}\}))(1+\log (d))\}  \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\] where in this case $\vp = p$. This can be rewritten as 
\[
T(\varepsilon^{-1},d,\vp) = e^p[\max\{1,\varepsilon^{-1}\}]^{p(1+\log d)}d^p
\]
Observe that the function $T$ is a function of $\max\{1,\varepsilon^{-1}\}$ and $d$ which are, polynomial or not exponential in powers of $\max\{1,\varepsilon^{-1}\}$ and $d$. Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for quasi-polynomial tractability in the algebraic case
\[
\sup_{d\in\mathbb{N}}\sum_{i=Lp\,e^pd^{p}} \left(\min\{1,\lambda_{i,d}\} \right)^{p(1+\ln d)}e^{-p}d^{-p} < \infty \qquad \text{for some} p \geq 0, L_p \in \naturals.
\]
\end{example}
\begin{example}[Exponential quasi-polynomial tractability]

\noindent Suppose that the tractability function, $T$, is defined by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log (d))\}  \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\] where in this case $\vp = p$. This can be rewritten as 
\[
T(\varepsilon^{-1},d,\vp) = e^p[\max\{1,\log(1+\varepsilon^{-1}\})]^{p(1+\log d)}d^p
\]
Observe that the function $T$ is a function of $\max\{1,\log(1+\varepsilon^{-1})\}$ and $d$. Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for quasi-polynomial tractability in the exponential case
\[
\sup_{d\in\mathbb{N}}\sum_{i=Lp\,e^pd^{p}} \left(\min\{1,\log(1+\lambda_{i,d})\} \right)^{p(1+\ln d)}e^{-p}d^{-p} < \infty
\]
\end{example}

\kachi{In the following we introduce two specific cases of tractability which depend on the form of the function $T$.
\begin{example}[(Strong) Separable tractability]
Suppose that
\[
 T(\varepsilon^{-1},d,\vp)= \phi(\varepsilon^{-1},p)\, \psi(d,q)
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2.
\]
with non-decreasing functions $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$ and where $\vp = (p,q)$.

We say that a problem is defined as $T$-separable tractable with parameter $\vp$ iff there exists a positive constant $C_{\vp}$ which does not depend on $\varepsilon$ and $d$ such that
\[
\comp(\varepsilon,d) \leq  C_{\vp}\,T(\varepsilon^{-1},d,\vp)\qquad \forall \varepsilon >0, \, d\in\naturals,
\]
that is
\[
\comp(\varepsilon,d) \leq \phi(\varepsilon^{-1},p)\, \psi(d,q) \qquad \forall \varepsilon >0, \, d\in\naturals.
\]
A problem is strongly $T$-separable tractable with parameter $\vp$ iff the complexity is independent of $d$, that is, there exists a positive constant $C_{\vp}$ which again does not depend on $\varepsilon$ and $d$ such that
\[
\comp(\varepsilon,d) \leq C_{\vp}T(\varepsilon^{-1}, 1,\vp) \qquad \forall \varepsilon >0, \, d\in\naturals,
\] where $\vp = p$. We have
\[
\comp(\varepsilon,d) \leq \phi(\varepsilon^{-1},p)\qquad \forall \varepsilon >0, \, d\in\naturals.
\]
%In addition, we assume the following conditions as in (\ref{eq:ptauassume}). That is, there exists a $K_{(p,q),\tau}$ depending on $p,q$ and $\tau$, but independent of $\varepsilon$ and $d$, such that
%\begin{equation} \label{eq:tractiff}    \sup_{d \in \naturals} \frac{1}{\psi(d,q)} \sum_{i = C(p,q)}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p))} < \infty.\end{equation}
Furthermore, we assume that
\[
\lim_{\varepsilon\rightarrow0}T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d\in\naturals, \vp \in[0,\infty)^2.
\]
Also since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit
\[
T(0,d,\vp):= \lim_{\varepsilon \rightarrow \infty}T(\varepsilon^{-1},d,\vp)\geq T(0,1,\textbf{0})>0
\] 
%Furthermore, we assume the following technical conditions. There exists a $K_{(p,q),\tau}$ depending on $p,q$ and $\tau$, but independent of $\varepsilon$ and $d$, such that
\begin{align*}
     [\phi(\varepsilon^{-1},p)]^{\tau} &\le K_{(p,1),\tau} \phi (\varepsilon^{-1},\tau p)\\
      [\psi(d,q)]^{\tau} &\le K_{(1,q),\tau} \psi (d,\tau q) \qquad \forall \, \varepsilon \in (0,\infty),\, d\in \naturals,\, p,q\geq 0,\, \tau >0.
\end{align*}
%\begin{align*}
 %  [T(\varepsilon^{-1},d,\vp]^\tau \leq K_{\vp,\tau}T(\varepsilon^{-1},d,\tau\vp) \quad \forall \varepsilon \in (0,\infty),\, d\in \mathbb{N},\, \vp\in[0,\infty)^2, \tau >0\end{align*}
Thus, Theorem \ref{thm_main_strong_tract2} yields the following equivalent condition for strongly separable tractability
\begin{equation*} \label{eq:strong_tractiff}
     \sup_{d \in \naturals} \sum_{i = L_p}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty.
\end{equation*}
Theorem \ref{thm_main_tract2} yields the following equivalent condition for 
separable tractability
\[
 \sup_{d \in \naturals} \frac{1}{\psi(d,q)} \sum_{i = L_{(p,q)\psi(d,q)}}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty.
\]
%As the name implies, the roles of $\varepsilon^{-1}$ and $d$ are separated. 
\end{example}
\begin{example}[Non-separable tractability]
    Let the tractability function $T$, be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(\phi(\varepsilon^{-1})\, \psi(d))\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp\in [0,\infty).
\]  with non-decreasing functions $\phi : (0,\infty) \to (0,\infty)$ and $\psi : \naturals \to (0,\infty)$ and $\vp = p$. Furthermore, we assume the following technical conditions. There exists a $K_{(p,q),\tau}$ depending on $p,q$ and $\tau$, but independent of $\varepsilon$ and $d$, such that
\begin{align*}
     [\phi(\varepsilon^{-1},p)]^{\tau} &\le K_{(p,1),\tau} \phi (\varepsilon^{-1},\tau p)\\
      [\psi(d,q)]^{\tau} &\le K_{(1,q),\tau} \psi (d,\tau q) \qquad \forall \, \varepsilon \in (0,\infty),\, d\in \naturals,\, p,q\geq 0,\, \tau >0.
\end{align*}
\end{example}
}

@@

Suppose that the domain of interesting $(\varepsilon^{-1},d)$ is not all of $(0,\infty \times \naturals$ but given by 
\begin{equation}
    \Omega  = \{(\varepsilon^{-1},d) : \varepsilon > E(d) \ge 0, \ d \in \naturals \},
\end{equation}
where $E$ is a non-increasing function.  A problem is
$T$-tractable with parameter $\vp$ \emph{on the restricted domain $\Omega$} iff the complexity is independent of the dimension of the problem, that is, there exists a positive constant $C_{\vp}$, again independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:rest_strong_tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},1,\vp) \qquad \forall(\varepsilon^{-1},d) \in \Omega.
\end{equation}
Define $M(d) = \max\{n : \lambda_{n+1,d} \ge E(d)\}$, i.e., 
\[
\lambda_{1,d}\ge  \cdots \ge \lambda_{M(d)+1,d} \ge E(d) > \lambda_{M(d)+2,d}, \qquad
(\lambda_{1,d}^{-1},d), \ldots, (\lambda^{-1}_{M(d)+1,d},d)  \in \Omega.
\]

>>>>

\newcommand{\tpmod}[1]{{\;(\operatorname{mod}\;#1)}}
\newcommand{\tI}{\widetilde{I}}

@@

The question of tractability is one of determining how fast the complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  We want to bound the complexity in terms of a simple function of $\varepsilon^{-1}$ and $d$ and identify simple conditions on the singular values that are equivalent to this bound holding.

In this article, we investigate such a function that we refer to as the generalized tractability function, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty),
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we we define our approximation problem to be tractable if 
$\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant, $C_{\vp}$, depending only on the parameter $\vp$.  In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase mildly with increasing $\varepsilon^{-1}$ and/or $d$, however most of our theorems do not require such an assumption.

@@

In Section \ref{sec:spt}, we give the definitions of the considered tractability notions, in particular (strong) tractability. We consider the strong tractability when the information complexity is independent of the dimension, $d$. We then turn to tractability  when the information complexity is polynomial in $d$ and $\varepsilon^{-1}$. In Section \ref{sec:examples} we provide examples of strong(tractability) functions for both the algebraic and exponential cases. We also consider other notions of tractability such as quasi-polynomial tractability, and a generalized notion of weak tractability. In the case of the generalized weak tractability, we give the necessary and sufficient conditions on the notion of exponential tractability in Section \ref{sec:subh}.

@@

\subsection{Notation and fundamental definitions} \label{sec:not}
Let $T$ be a function given in \eqref{eq:Tspec}.
A tractability function, $T$ is a function of simple form that provides an upper bound on the computational complexity of a problem.  Some examples of tractability functions commonly considered are provided in Table \ref{tab:commonT}.

\begin{table} 
    \caption{Common forms of the tractability function, $T$}
{\small
\begin{equation*}
	\begin{array}{r@{\quad}c@{\quad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		\text{strong polynomial algebraic} & \max(1,\varepsilon^{-p}) & 1\\
		\text{polynomial algebraic} & \max(1,\varepsilon^{-p})d^{q} & d^{q}\\
		\text{strong polynomial exponential} &  [\max(1,\log(1 + \varepsilon^{-1})]^p & 1\\
		\text{polynomial exponential } &
		[\max(1,\log(1 + \varepsilon^{-1}))]^p  d^{q} & d^{q} \\
        \text{quasi-polynomial algebraic} & 
        \exp\{p(1+\log(\max(1,\varepsilon^{-1}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}\\
        \text{quasi-polynomial exponential} & 
        \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}
\label{tab:commonT}
\end{table}

@@

For this definition of tractability to make sense we assume that
\begin{subequations} \label{eq:Tconditions}
\begin{equation} \label{eq:Tnondecreasing}
	T \text{ is non-decreasing in all variables,}
\end{equation}
which implies that the problem is expected to be no easier by decreasing the tolerance, $\varepsilon$, or increasing the dimension, $d$. Furthermore, increasing $\vp$ does not decrease $T(\cdot, \cdot, \vp)$.  Since there are an infinite number of positive eigenvalues, it also makes sense to assume that
\begin{equation}
	\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d \in \naturals, \ \vp \in [0,\infty)^s.
\end{equation}
Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit, which we assume will always be positive:
\begin{equation}\label{eq:T0}	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp) \ge T(0,1,\vzero) > 0.
\end{equation}
\end{subequations}
Finally, there is a technical assumption required for our analysis.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp),   \quad \forall \varepsilon \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in [1,\infty).
\end{equation}

Certain common notions of tractability  covered by the general situation described above are provided in Table \ref{tab:commonT}.  Here, $\vp = (p,q)$ in some cases and $\vp$ is a scalar in other cases. A general reference
for results on algebraic tractability is \cite{NW08}, for exponential tractability we refer to \cite{HKW19} and \cite{KW19}.

@@

For $\varepsilon$ taking on all positive values, $i(\varepsilon,d,\vp)$ takes on all values greater than or equal to $i(\infty,d,\vp)$, so
\[
i\le C_{\vp}\, T(\varepsilon^{-1},d,\vp)+1 \le C_{\vp} T(\lambda_{i,d}^{-1},d,\vp)+1, \qquad \forall i \ge i(\infty,d,\vp),  \ d \in \naturals.
\]
This implies via our technical assumption \eqref{eq:ptauassume} that for any $\tau > 1$,
\begin{gather*}
	K_{\vp,\tau}\,T (\lambda_{i,d}^{-1},d,\tau \vp) \ge
	[T(\lambda_{i,d}^{-1},d, \vp)]^\tau
	\ge
	\left[\frac{(i-1)}{C_{\vp}}\right]^\tau \qquad \forall i \ge i(\infty,d,\vp),  \ d \in \naturals, \\
	 \frac{1}{T (\lambda_{i,d}^{-1},d,\tau \vp)} \le
	\frac{K_{\vp,\tau}\, C_{\vp}^\tau}{(i-1)^\tau}\qquad \forall i \ge \max\{i(\infty,d,\vp),2\},  \ d \in \naturals.
\end{gather*}

@@

\begin{example}[Algebraic polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-p}\}\, d^q
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where in this case $\vp = (p,q)$.
Then Theorem \ref{thm_main_tract2} yields the  equivalent condition for algebraic polynomial tractability:
\[
 \sup_{d\in\naturals}\, d^{-q} \sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty (\min\{1,\lambda_{i,d}^p\})  < \infty \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\]
Furthermore, the set of optimal $(p,q)$ is a subset of
\begin{multline*}
\{(p^*,q^*) : \text{the above condition holds for all }(p,q) \in (p^*,\infty) \times (q^*,\infty) \\
\text{but not for any }(p,q) \in [0,p^*/2) \times [0,q^*/2)\}.
\end{multline*}
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NW08} \fred{same theorem as above??, what about optimality in \cite[Theorem 5.1]{NW08}}.
\end{example}

\begin{example}[Strong exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \left[\max\{1,\log(1+\varepsilon^{-1})\}\right]^p, \quad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]
where $\vp$ is a scalar. Then Theorem \ref{thm_main_strong_tract2} yields the equivalent condition for strong exponential polynomial tractability \[
 \sup_{d\in\naturals} \sum_{i=L_{p} }^\infty (\min\{1,\log(1+\lambda_{i,d}\})^{p}  < \infty,\quad  \text{ for some } p \ge 0, L_{p} \in \naturals
\] In this case the optimal $p$ is the infimum of all $p$ for which the above condition holds.
\end{example}
\begin{example}[Exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)=\left[\max\{1,\log(1+\varepsilon^{-1}\})\right]^p\, d^q, \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where $\vp = (p,q)$. Then Theorem \ref{thm_main_tract2} yields  the equivalent condition for exponential polynomial tractability
\[
 \sup_{d\in\naturals}\, d^{-q} \sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty [\min\{1,\log(1+\lambda_{i,d}\})]^p < \infty  \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\]
\end{example}

\begin{example}[(Strong) Separable tractability]
Suppose that the tractability function is defined by
\[
 T(\varepsilon^{-1},d,\vp)= \phi(\varepsilon^{-1},p)\, \psi(d,q)
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2.
\]
with non-decreasing functions $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$ and where $\vp = (p,q)$, hence $T$ is non-decreasing in all variables.

We say that a problem is defined as $T$-separable tractable with parameter $\vp$ iff there exists a positive constant $C_{\vp}$ which does not depend on $\varepsilon$ and $d$ such that
\[
\comp(\varepsilon,d) \leq  C_{\vp}\,\phi(\varepsilon^{-1},p)\, \psi(d,q)\qquad \forall \varepsilon >0, \, d\in\naturals,
\] where $\vp = (p,q)$.
A problem is strongly $T$-separable tractable with parameter $\vp$ iff the complexity is independent of $d$, that is, there exists a positive constant $C_{\vp}$ which again does not depend on $\varepsilon$ and $d$ such that
\[
\comp(\varepsilon,d) \leq C_{\vp}\phi(\varepsilon^{-1},p)\qquad \forall \varepsilon >0, \, d\in\naturals,
\] where $\vp = p$ is a scalar. 
Furthermore, we assume the following as in \eqref{eq:Tconditions} and \eqref{eq:ptauassume}. We assume that
\[
\lim_{\varepsilon\rightarrow0}\phi(\varepsilon^{-1},p) = \infty \qquad \forall d\in\naturals, p >0.
\] since there are an infinite number of positive eigenvalues. Furthermore, we have that $\phi(\cdot,p)$ is non-decreasing hence we define the following limit
\[ \lim_{\varepsilon \rightarrow \infty}\phi(\varepsilon^{-1},p)\geq 1.
\] 
Additionally, we have the following technical assumptions. Suppose that, for any real $\tau_1, \tau_2>0$ and any $p,q>0$, we have that there exists $\hat{K}_{p,\tau_1}$ and $\hat{K}_{q\tau_2}$ such that
\begin{align*}
     [\phi(\varepsilon^{-1},p)]^{\tau_1} &\le \hat{K}_{p,\tau_1} \phi (\varepsilon^{-1},\tau_1 p)\\
      [\psi(d,q)]^{\tau_2} &\le \hat{K}_{q,\tau_2} \psi (d,\tau_2 q) \qquad \forall \, \varepsilon \in (0,\infty),\, d\in \naturals.
\end{align*} where $\hat{K}_{p,\tau_1}>0$ is a constant that may depend on $p$ and $\tau_1$, but is independent of $\varepsilon$, and where $\hat{K}_{q,\tau_2}>0$ is a constant that may depend on $q$ and $\tau_2$, but is independent of $d$.

Thus, Theorem \ref{thm_main_strong_tract2} yields the following equivalent condition for strongly $T$-separable tractability
\begin{equation*} \label{eq:strong_tractiff}
     \sup_{d \in \naturals} \sum_{i = L_p}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty.
\end{equation*}
Theorem \ref{thm_main_tract2} yields the following equivalent condition for 
separable tractability
\[
 \sup_{d \in \naturals} \frac{1}{\psi(d,q)} \sum_{i = \lceil L_{(p,q)\psi(d,q)}\rceil}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty,
\] for some $p,q >0$ and $L_{(p,q)}\in\naturals$.
\end{example}

\begin{example}[Non-separable tractability]
    Suppose we define the tractability function $T$ by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(\phi(\varepsilon^{-1})\, \psi(d))\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp\in [0,\infty).
\]  with non-decreasing functions $\phi : (0,\infty) \to (0,\infty)$ and $\psi : \naturals \to (0,\infty)$ and $\vp = p$. 
We say that a problem is non-separable tractable with parameter $\vp$ iff there exists a positive $C_{\vp}$ such that

\[\comp(\varepsilon,d) \leq  C_{\vp}\exp\{p(\phi(\varepsilon^{-1})\psi(d))\}.\]

Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for non-separable tractability
\[
\sup_{d\in\mathbb{N}}\frac{1}{\psi(d)}\sum_{i=Lp\,\psi(d)} \phi(\lambda_{i,d})\quad \text{for some } p \geq 0, L_p \in \naturals.
\]
We have the following special cases of non-separable tractability. \end{example}
\begin{example}[Algebraic quasi-polynomial tractability]

@@

\noindent Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(1+\log(\max\{1,\varepsilon^{-1}\}))(1+\log (d))\}  \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\] where in this case $\vp = p$. This can be rewritten as 
\[
T(\varepsilon^{-1},d,\vp) = e^p[\max\{1,\varepsilon^{-1}\}]^{p(1+\log d)}d^p
\]
Observe that the function $T$ is a function of $\max\{1,\varepsilon^{-1}\}$ and $d$ which are, polynomial or not exponential in powers of $\max\{1,\varepsilon^{-1}\}$ and $d$. Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for quasi-polynomial tractability in the algebraic case
\[
\sup_{d\in\mathbb{N}}\,d^{-p}\sum_{i=Lp\,d^{p}} \left(\min\{1,\lambda_{i,d}\} \right)^{p(1+\ln d)} < \infty \qquad \text{for some } p \geq 0, L_p \in \naturals.\]
\end{example}
\begin{example}[Exponential quasi-polynomial tractability]

\noindent Suppose that the tractability function, $T$, is defined by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log (d))\}  \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\] where in this case $\vp = p$. This can be rewritten as 
\[
T(\varepsilon^{-1},d,\vp) = e^p[\max\{1,\log(1+\varepsilon^{-1}\})]^{p(1+\log d)}d^p.
\]
Observe that the function $T$ is a function of $\max\{1,\log(1+\varepsilon^{-1})\}$ and $d$. Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for quasi-polynomial tractability in the exponential case
\[
\sup_{d\in\mathbb{N}}\,d^{-p}\sum_{i=Lp\,d^{p}} \left(\min\{1,\log(1+\lambda_{i,d})\} \right)^{p(1+\ln d)}e^{-p} < \infty \quad \text{for some } p \geq 0, L_p \in \naturals.
\]
\end{example}

%In the following we introduce two specific cases of tractability which depend on the form of the function $T$ (see \cite{}).

@@

Suppose that the domain of interest $(\varepsilon^{-1},d)$ is not all of $(0,\infty) \times \naturals$ but given by 
\begin{equation}
    \Omega  = \{(\varepsilon^{-1},d) : \varepsilon \peter{> E(d) \ge 0}, \ d \in \naturals \},
\end{equation}
%where $E$ is a non-increasing \peter{[{is this a necessary assumption?}]} function.  
A problem is \peter{strongly}
$T$-tractable with parameter $\vp$ \emph{on the restricted domain $\Omega$} iff the complexity is independent of the dimension of the problem, that is, there exists a positive constant $C_{\vp}$, again independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:rest_strong_tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},1,\vp) \qquad \forall(\varepsilon^{-1},d) \in \Omega.
\end{equation}
\peter{[I am wondering whether this makes sense if $E$ depends on $d$? Do Gnewuch and Wo\'{z}niakowski allow dependence of $E$ on $d$?]}
